\name: Deterministic and Coherent Implicits

### 2018-06-19 jfp/src/Translation.tex
changes

%-------------------------------------------------------------------------------
\subsection{Type-Directed Translation}
Figure~\ref{fig:type} presents the translation rules that convert $\name$
expressions into ones of System~F. 
The gray parts of the figure extend the type system with the necessary
information for the translation.

@@

Moreover, we can express three key properties of Figure~\ref{fig:resolution2}'s
definition of resolution in terms of the generated evidence.
Firstly, the deterministic version of resolution is a sound variation on the original ambiguous resolution.
\begin{theorem}[Soundness]
Figure~\ref{fig:resolution2}'s definition of resolution is sound (but
incomplete) with respect to Figure~\ref{fig:resolution1}'s definition.
\[\forall \tenv, \rulet, E: \quad\quad \dres{\tenv}{\rulet}{E} \quad\Rightarrow\quad \ares{\tenv}{\rulet}{E} \]
\end{theorem}
Secondly, the deterministic resolution guarantees is effectively deterministic:
\begin{theorem}[Determinacy]
The generated evidence of resolution is uniquely determined.
\[\forall \tenv, \rulet, E_1, E_2: \quad\quad 
     \dres{\tenv}{\rulet}{E_1} ~\wedge~ 
     \dres{\tenv}{\rulet}{E_2} \quad\Rightarrow\quad E_1 = E_2 \]
\end{theorem}
It follows immediately that deterministic resolution is also coherent,
which is formulated as follows for our elaboration-based setting.
\begin{corollary}[Coherence]
\[\forall \tenv, \rulet, E_1, E_2: \quad\quad 
     \dres{\tenv}{\rulet}{E_1} ~\wedge~ 
     \dres{\tenv}{\rulet}{E_2} \quad\Rightarrow\quad 
     \fctx{\etrans{\tenv}}{E_1}{E_2}{\ttrans[\rulet]} \]
\end{corollary}
Here the judgment $\fctx{\tenv}{E_1}{E_2}{T}$
denotes \emph{contextual equivalence}, which captures that
$E_1$ and $E_2$ behave the same in any well-typed program context. We do not give the
formal definition here as it is not required to prove the corollary; we only
need to know that it is an equivalence relation, and, more spefically, that
it is reflexive, i.e., any well-typed System~F term is contextually equivalent
to itself.

>>>>

%-------------------------------------------------------------------------------
\subsection{Type-Directed Translation}
Figure~\ref{fig:type} presents the translation rules that convert $\name$
expressions into System~F expressions. 
The gray parts of the figure extend the type system with the necessary
information for the translation.

@@

Moreover, we can express three key properties of Figure~\ref{fig:resolution2}'s
definition of resolution in terms of the generated evidence.
Firstly, the deterministic version of resolution is a sound variation on the original ambiguous resolution.
\begin{theorem}[Soundness]
Figure~\ref{fig:resolution2}'s definition of resolution is sound (but
incomplete) with respect to Figure~\ref{fig:resolution1}'s definition.
\[\forall \tenv, \rulet, E: \quad\quad \dres{\tenv}{\rulet}{E} \quad\Rightarrow\quad \ares{\tenv}{\rulet}{E} \]
\end{theorem}
Secondly, the deterministic resolution is effectively deterministic:
\begin{theorem}[Determinacy]
The generated evidence of resolution is uniquely determined.
\[\forall \tenv, \rulet, E_1, E_2: \quad\quad 
     \dres{\tenv}{\rulet}{E_1} ~\wedge~ 
     \dres{\tenv}{\rulet}{E_2} \quad\Rightarrow\quad E_1 = E_2 \]
\end{theorem}
It follows immediately that deterministic resolution is also coherent,
which is formulated as follows for our elaboration-based setting.
\begin{corollary}[Coherence]
\[\forall \tenv, \rulet, E_1, E_2: \quad\quad 
     \dres{\tenv}{\rulet}{E_1} ~\wedge~ 
     \dres{\tenv}{\rulet}{E_2} \quad\Rightarrow\quad 
     \fctx{\etrans{\tenv}}{E_1}{E_2}{\ttrans[\rulet]} \]
\end{corollary}
Here the judgment $\fctx{\tenv}{E_1}{E_2}{T}$
denotes \emph{contextual equivalence}~\cite{reynolds?}, which captures that
$E_1$ and $E_2$ behave the same in any well-typed program context. We do not give the
formal definition here as it is not required to prove the corollary; we only
need to know that it is an equivalence relation, and, more spefically, that
it is reflexive, i.e., any well-typed System~F term is contextually equivalent
to itself. 
\bruno{Should we not mention that contextual equivalence for System F
  is well-studied, thus interested readers can consult the definition
  in other works?}
